A Horrocks-type theorem for even orthogonal $K_2$

### 2020-04-09 dl-paper.tex
Beautify 4.6

\section{Preliminaries}
If $x$ and $y$ are elements of a group $G$ then $[x, y]$ denotes the left-normed commutator $xyx^{-1}y^{-1}$.
We denote by $x^y$ (resp. ${}^y\!x$) the element $y^{-1}xy$ (resp. $yxy^{-1}$). 
Throughout the paper we make use of the following commutator identities:
\begin{align}
 [x, yz]^y =& [y^{-1}, x] \cdot [x, z], \label{rel43} \\ 
 \label{eq:H1ii} [xy, z] =& {}^x[y, z] \cdot [x,z], \\
 [[x, y], z] =& \left([x^{-1}, [ y^{-1}, z]] ^ {y^{-1}} \cdot [y, [ z^{-1}, x^{-1}]] ^ {z^{-1}} \right)^{x^{-1}}. \label{HW-variant} \end{align}
Recall also from~\cite[Lemma~3.1.1]{RS76} that
\begin{equation} \label{eq:H1iii} [x,z] = 1 \text{ implies } [x, [y,z]] = [[x,y],{}^yz]. \end{equation}

@@

In this paper we will be mostly interested in the case when the Dynkin diagram of $\Phi$ is {\it simply-laced}, i.\,e. does not contain double bonds. In this case the defining relations of $\St(\Phi, R)$ have the following simpler form:
\begin{align}
&x_{\alpha}(a)\cdot x_{\alpha}(b)=x_{\alpha}(a+b), \tag{R1} \label{Steinberg-additivity}\\
&[x_{\alpha}(a),\,x_{\beta}(b)]=x_{\alpha+\beta}(N_{\alpha\beta} \cdot ab),\text{ for }\alpha+\beta\in\Phi, \tag{R2} \label{Chevalley-CCF1} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]=1,\text{ for }\alpha+\beta\not\in\Phi\cup0. \tag{R3} \label{Chevalley-CCF2}
\end{align}
In the above formulae $a, b \in R$ and the integers $N_{\alpha, \beta} = N_{\alpha, \beta, 1, 1} = \pm 1$ are the structure constants of the Lie algebra of type $\Phi$. Although there is still some degree of freedom in their choice, they always must satisfy the relations, indicated in the following lemma (cf. \cite[\S~14]{VP}).
\begin{lemma} Suppose $\Phi$ is simply laced and $\alpha, \beta$ are roots of $\Phi$ such that $\alpha+\beta\in \Phi$, then the following identities hold for the structure constants:
\begin{equation} \label{eq:simplest} N_{\alpha, \beta} = -N_{\beta,\alpha} = - N_{-\alpha, -\beta} = N_{\beta, -\alpha-\beta} = N_{-\alpha-\beta, \alpha}. \end{equation}
If, moreover, $\gamma \in \Phi$ is such that $\alpha,\beta,\gamma$ form a basis of a root subsystem of type $\rA_3$ then one has
\begin{equation} \label{eq:cocycle} N_{\beta,\gamma} \cdot N_{\alpha, \beta+\gamma} = N_{\alpha+\beta, \gamma} \cdot N_{\alpha, \beta}. \end{equation} \end{lemma}
In our calculations below we will be using identities~\eqref{eq:simplest} without further reference.

@@

Recall from~\cite[Lemma~5.2]{Ma69} that the following relations hold for semisimple root elements:
\begin{align} \label{eq:conj-h-x} {}^{h_\alpha(t)}\!x_\beta(u) & = x_\beta(t^{\langle \beta,  \alpha \rangle}u), \\
              \label{eq:conj-h-h} {}^{h_\alpha(t)}\!h_\beta(u) & = h_\beta(t^{\langle \beta, \alpha \rangle} \cdot u) \cdot h_\beta(t^{\langle \beta,  \alpha \rangle})^{-1}, \\
              \label{eq:h-inv}    h_\alpha(t)^{-1}             & = h_{-\alpha}(t). \end{align}
Here $\langle \beta, \alpha \rangle$ denotes the integer $\tfrac{2(\beta, \alpha)}{(\alpha, \alpha)}$ (we denote by $(-,-)$ the inner product on the euclidean space $\mathbb{R}^\ell$ containing $\Phi$).

@@

\begin{lemma} \label{Crels} Let $\Phi$ be a simply-laced root system.
The elements $c_\alpha(s, t)$ satisfy the following relations for all $s\in I,\ t\in J,\ \xi\in R$:
 \begin{enumerate}
 \item \label{C1} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = x_{\alpha}(- st\xi) \cdot x_{\alpha+\beta}(N_{\alpha,\beta}\cdot s^2t\xi)$ if $\alpha+\beta \in \Phi$;
 \item \label{C2} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = x_{\alpha}(st\xi + s^2t^2\xi) \cdot x_{\alpha-\beta}(N_{-\alpha, \beta}\cdot st^2\xi)$ if $\alpha-\beta \in \Phi$;  
 \item \label{C3} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = 1$ if $\alpha \perp \beta$;  
 \item \label{C4} If $\alpha+\beta\in\Phi$ then
  \begin{equation} \label{eq:C4} c_{\alpha+\beta}(s, t\xi) = [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot c_{\alpha}(\epsilon s\xi, -\epsilon t)^{-1} \cdot x_{-\beta}(-st\xi^2),\end{equation}
  where $\epsilon = N_{\alpha,\beta}$.
 \end{enumerate}
\end{lemma}
\begin{proof}
The first assertion follows from the following computation:
\begin{multline} \nonumber [c_{\beta}(s, t),\ x_{\alpha}(\xi)]  = [x_{\beta}(s),\ x_{-\beta}(t)] \cdot [x_{-\beta}(t),\ x_{\beta}(s)\cdot x_{\alpha + \beta}(N_{\alpha, \beta} \cdot s\xi)] = \\
 = {}^{x_\beta(s)}\![x_{-\beta}(t),\ x_{\alpha+\beta}(N_{\alpha, \beta} \cdot s\xi)] = x_{\alpha}(- st\xi) \cdot x_{\alpha+\beta}(N_{\alpha,\beta} \cdot s^2t\xi). \end{multline}

The second assertion follows directly from~\eqref{HW-variant} and Chevalley commutator identities:
\begin{multline} \nonumber
[[x_\beta(s),\ x_{-\beta}(t)],\ x_{\alpha}(\xi)] = \\
= \left([x_\beta(-s),\ [x_{-\beta}(-t),\ x_{\alpha}(\xi)]] ^ {x_{-\beta}(-t)} \cdot [x_{-\beta}(t),\ [ x_{\alpha}(-\xi),\ x_\beta(-s)]] ^ {x_{\alpha}(-\xi)}\right)^{x_\beta(-s)} = \\
= x_{\alpha}(-N_{\beta, \alpha-\beta} N_{\alpha,-\beta} \cdot s t \xi) ^ {x_{-\beta}(-t) \cdot x_\beta(-s)} = x_{\alpha}(s t \xi) ^ {x_{-\beta}(-t) \cdot x_\beta(-s)} = \\
= x_{\alpha}(st\xi) \cdot x_{\alpha-\beta}(-N_{\alpha,-\beta}\cdot st^2\xi)^{x_\beta(-s)} = x_{\alpha}(st\xi + s^2t^2 \xi) \cdot x_{\alpha-\beta}(N_{-\alpha,\beta}\cdot st^2\xi). \end{multline}

Finally, the last assertion can be verified via the following direct computation:
\begin{align*}
[x_{\alpha+\beta}(s),\ x_{-\alpha-\beta}(t\xi)] = [x_{\alpha+\beta}(s),\ [x_{-\alpha}(-\epsilon t),\ x_{-\beta}(\xi)]]&\ \text{by~\eqref{Chevalley-CCF1}}\\
\begin{split}
    =\bigl( [[x_{\alpha+\beta}(-s),\ x_{-\alpha}(\epsilon t)],\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s)} \cdot \hspace{70pt} \\
        \cdot [[ x_{-\beta}(-\xi),\ x_{\alpha+\beta}(s)],\ x_{-\alpha}(\epsilon t)]^{x_{-\beta}(-\xi)}\bigr)^{x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{HW-variant}}
\end{split}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha}(\epsilon t)]^{x_{-\beta}(-\xi)x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{Chevalley-CCF1}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha}(\epsilon t) x_{-\alpha-\beta}(t\xi)]^{x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{Chevalley-CCF1},\eqref{Chevalley-CCF2}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot [x_{-\alpha}(-\epsilon t),\ x_{\alpha}(\epsilon s\xi)] \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha-\beta}(t\xi)] &\ \text{by~\eqref{rel43}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot c_{\alpha}(\epsilon s\xi, -\epsilon t)^{-1} \cdot x_{-\beta}(-st\xi^2)  &\ \text{by~\eqref{Chevalley-CCF1}}. \qedhere
\end{align*}
\end{proof}

@@

\begin{lemma}\label{P0_normal} The subgroup $\widetilde{P}_{\alpha, M}(0)$ is normal in $\widetilde{P}_{\alpha, M}(*)$. Moreover, there is a short exact sequence of groups, which is split by the map $m \mapsto x_{-\alpha}(mX)$ (we denote by $(M, +)$ the additive group of the ideal $M$):
\[\xymatrix{1 \ar[r] & \widetilde{P}_{\alpha,M}(0) \ar[r] & \widetilde{P}_{\alpha,M}(*) \ar@{->>}[r]^{p_\alpha} & (M, +) \ar@<5pt>@{-->}[l] \ar[r] & 1}.\] \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ to every generator $g$ of $\widetilde{P}_\alpha(0)$ listed in~\cref{defP0} belongs to $\widetilde{P}_\alpha(0)$.
The assertion is obvious for the generators of type P3 and P4.

>>>>

\section{Preliminaries}
If $x$ and $y$ are elements of a group $G$ then $[x, y]$ denotes the left-normed commutator $xyx^{-1}y^{-1}$.
We denote by $x^y$ (resp. ${}^y\!x$) the element $y^{-1}xy$ (resp. $yxy^{-1}$). 
Throughout the paper we make use of the following commutator identities:
\begin{align}
 [x, yz]^y =& [y^{-1}, x] \cdot [x, z], \label{rel43} \\ 
 \label{eq:H1ii} [xy, z] =& {}^x[y, z] \cdot [x,z], \\
 \label{eq:H1ii-2}[x, yz] =& [x, y] \cdot {}^{y}\![x, z], \\
 [[x, y], z] =& \left([x^{-1}, [ y^{-1}, z]] ^ {y^{-1}} \cdot [y, [ z^{-1}, x^{-1}]] ^ {z^{-1}} \right)^{x^{-1}}. \label{HW-variant} \end{align}
Recall also from~\cite[Lemma~3.1.1]{RS76} that
\begin{equation} \label{eq:H1iii} [x,z] = 1 \text{ implies } [x, [y,z]] = [[x,y],{}^yz]. \end{equation}

@@

In this paper we will be mostly interested in the case when the Dynkin diagram of $\Phi$ is {\it simply-laced}, i.\,e. does not contain double bonds. In this case the defining relations of $\St(\Phi, R)$ have the following simpler form:
\begin{align}
&x_{\alpha}(a)\cdot x_{\alpha}(b)=x_{\alpha}(a+b), \tag{R1} \label{Steinberg-additivity}\\
&[x_{\alpha}(a),\,x_{\beta}(b)]=x_{\alpha+\beta}(N_{\alpha\beta} \cdot ab),\text{ for }\alpha+\beta\in\Phi, \tag{R2} \label{Chevalley-CCF1} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]=1,\text{ for }\alpha+\beta\not\in\Phi\cup0. \tag{R3} \label{Chevalley-CCF2}
\end{align}
In the above formulae $a, b \in R$ and the integers $N_{\alpha, \beta} = N_{\alpha, \beta, 1, 1} = \pm 1$ are the structure constants of the Lie algebra of type $\Phi$. Although there is still some degree of freedom in their choice, they always must satisfy the relations, indicated in the following lemma (cf. \cite[\S~14]{VP}).
\begin{lemma} Suppose $\Phi$ is simply laced and $\alpha, \beta$ are roots of $\Phi$ such that $\alpha+\beta\in \Phi$, then the following identities hold for the structure constants:
\begin{equation} \label{eq:simplest} N_{\alpha, \beta} = -N_{\beta,\alpha} = - N_{-\alpha, -\beta} = N_{\beta, -\alpha-\beta} = N_{-\alpha-\beta, \alpha}. \end{equation}
If, moreover, $\gamma \in \Phi$ is such that $\alpha,\beta,\gamma$ form a basis of a root subsystem of type $\rA_3$ then one has
\begin{equation} \label{eq:cocycle} N_{\beta,\gamma} \cdot N_{\alpha, \beta+\gamma} = N_{\alpha+\beta, \gamma} \cdot N_{\alpha, \beta}. \end{equation} \end{lemma}
In our calculations below we repeatedly use identities~\eqref{eq:simplest} without explicit reference.

@@

Recall from~\cite[Lemma~5.2]{Ma69} that the following relations hold for semisimple root elements:
\begin{align} \label{eq:conj-h-x} {}^{h_\alpha(t)}\!x_\beta(u) & = x_\beta(t^{\langle \beta,  \alpha \rangle}u), \\
              \label{eq:conj-h-h} {}^{h_\alpha(t)}\!h_\beta(u) & = h_\beta(t^{\langle \beta, \alpha \rangle} \cdot u) \cdot h_\beta(t^{\langle \beta,  \alpha \rangle})^{-1}, \\
              \label{eq:h-inv}    h_\alpha(t)^{-1}             & = h_{-\alpha}(t). \end{align}
In the above formulae $\langle \beta, \alpha \rangle$ denotes the integer $\tfrac{2(\beta, \alpha)}{(\alpha, \alpha)}$ (we denote by $(\text{-},\text{-})$ the inner product on the euclidean space $\mathbb{R}^\ell$ containing $\Phi$).

@@

\begin{lemma} \label{Crels} Let $\Phi$ be a simply-laced root system.
The elements $c_\alpha(s, t)$ satisfy the following relations for all $s\in I,\ t\in J,\ \xi\in R$:
 \begin{enumerate}
 \item \label{C1} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = x_{\alpha}(- st\xi) \cdot x_{\alpha+\beta}(N_{\alpha,\beta}\cdot s^2t\xi)$ if $\alpha+\beta \in \Phi$;
 \item \label{C2} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = x_{\alpha}(st\xi + s^2t^2\xi) \cdot x_{\alpha-\beta}(N_{-\alpha, \beta}\cdot st^2\xi)$ if $\alpha-\beta \in \Phi$;  
 \item \label{C3} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = 1$ if $\alpha \perp \beta$;  
 \item \label{C4} If $\alpha+\beta\in\Phi$ then
  \begin{equation} \label{eq:C4} c_{\alpha+\beta}(s, t\xi) = [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot c_{\alpha}(\epsilon s\xi, -\epsilon t)^{-1} \cdot x_{-\beta}(-st\xi^2),\end{equation}
  where $\epsilon = N_{\alpha,\beta}$.
 \end{enumerate}
\end{lemma}
\begin{proof}
The first two assertions can be verified by direct calculation:
\begin{align*} \nonumber [c_{\beta}(s, t),\ x_{\alpha}(\xi)] = [x_{\beta}(s),\ x_{-\beta}(t)] \cdot [x_{-\beta}(t),\ x_{\beta}(s)\cdot x_{\alpha + \beta}(N_{\alpha, \beta} \cdot s\xi)] &\text{ by~\eqref{Cdef}, \eqref{Chevalley-CCF1}, \eqref{Chevalley-CCF2}} \\
 = {}^{x_\beta(s)}\![x_{-\beta}(t),\ x_{\alpha+\beta}(N_{\alpha, \beta} \cdot s\xi)] &\text{ by~\eqref{eq:H1ii-2}} \\ = x_{\alpha}(- st\xi) \cdot x_{\alpha+\beta}(N_{\alpha,\beta} \cdot s^2t\xi) &\text{ by~\eqref{Chevalley-CCF1};} \end{align*}
\begin{align*} \nonumber
\begin{split}[[x_\beta(s),\ x_{-\beta}(t)],\ x_{\alpha}(\xi)] = \bigl([x_\beta(-s),\ [x_{-\beta}(-t),\ x_{\alpha}(\xi)]] ^ {x_{-\beta}(-t)} \cdot \hspace{50pt} \\ \cdot [x_{-\beta}(t),\ [ x_{\alpha}(-\xi),\ x_\beta(-s)]] ^ {x_{\alpha}(-\xi)}\bigr)^{x_\beta(-s)} & \text{ by~\eqref{HW-variant}}\end{split} \\
= x_{\alpha}(-N_{\beta, \alpha-\beta} N_{\alpha,-\beta} \cdot s t \xi) ^ {x_{-\beta}(-t) \cdot x_\beta(-s)} = x_{\alpha}(s t \xi) ^ {x_{-\beta}(-t) \cdot x_\beta(-s)} & \text{ by~\eqref{Chevalley-CCF1},\eqref{Chevalley-CCF2}}\\
= x_{\alpha}(st\xi) \cdot x_{\alpha-\beta}(-N_{\alpha,-\beta}\cdot st^2\xi)^{x_\beta(-s)} = x_{\alpha}(st\xi + s^2t^2 \xi) \cdot x_{\alpha-\beta}(N_{-\alpha,\beta}\cdot st^2\xi) & \text{ by~\eqref{Chevalley-CCF1}.} \end{align*}
The assertion~\eqref{C3} is obvious. Finally, last assertion can be obtained as follows:
\begin{align*}
[x_{\alpha+\beta}(s),\ x_{-\alpha-\beta}(t\xi)] = [x_{\alpha+\beta}(s),\ [x_{-\alpha}(-\epsilon t),\ x_{-\beta}(\xi)]]&\ \text{by~\eqref{Chevalley-CCF1}}\\
\begin{split}
    =\bigl( [[x_{\alpha+\beta}(-s),\ x_{-\alpha}(\epsilon t)],\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s)} \cdot \hspace{70pt} \\
        \cdot [[ x_{-\beta}(-\xi),\ x_{\alpha+\beta}(s)],\ x_{-\alpha}(\epsilon t)]^{x_{-\beta}(-\xi)}\bigr)^{x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{HW-variant}}
\end{split}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha}(\epsilon t)]^{x_{-\beta}(-\xi)x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{Chevalley-CCF1}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha}(\epsilon t) x_{-\alpha-\beta}(t\xi)]^{x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{Chevalley-CCF1},\eqref{Chevalley-CCF2}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot [x_{-\alpha}(-\epsilon t),\ x_{\alpha}(\epsilon s\xi)] \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha-\beta}(t\xi)] &\ \text{by~\eqref{rel43}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot c_{\alpha}(\epsilon s\xi, -\epsilon t)^{-1} \cdot x_{-\beta}(-st\xi^2)  &\ \text{by~\eqref{Chevalley-CCF1}}. \qedhere
\end{align*}

\end{proof}

@@

\begin{lemma}\label{P0_normal} The subgroup $\widetilde{P}_{\alpha, M}(0)$ is normal in $\widetilde{P}_{\alpha, M}(*)$. Moreover, there is a short exact sequence of groups, which is split by the map $m \mapsto x_{-\alpha}(mX)$ (we denote by $(M, +)$ the additive group of the ideal $M$):
\[\xymatrix{1 \ar[r] & \widetilde{P}_{\alpha,M}(0) \ar[r] & \widetilde{P}_{\alpha,M}(*) \ar@{->>}[r]^{p_\alpha} & (M, +) \ar@<5pt>@{-->}[l] \ar[r] & 1}.\] \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ of every generator $g$ of $\widetilde{P}_\alpha(0)$ listed in~\cref{defP0} belongs to $\widetilde{P}_\alpha(0)$.
The assertion is obvious for the generators of type P3 and P4.
